Joint training of classification model and similarity model for low-resource text classification in chatbot

### 2020-09-10 chatbot_classification/aaai_2021/paper.tex
update citation

  The   first  one,  \emph{text  classification},  includes  classic  long  text
  classification  and  short  text  classification  in  this  scenario.  Besides
  traditional  machine  learning  models  like  SVM,  boosting  tree,  many work
  \cite{wen2016network}  choose  neural  network  such  as  convolutional neural
  networks (CNNs) \cite{kim2014convolutional,zhang2015character,conneau2016very}
  and       long       short       term       memory       networks      (LSTMs)
  \cite{mousa2017contextual,liu2016recurrent}   to   accomplish   the   task  of
  extracting  semantic  feature from limited amount of words. Their common model
  structure  is  adding  a  softmax classifier as the final layer and mapping to
  labels.  Another  two  interesting lines of work, label-word joint models, and
  joint  NER  and  classification  are  discussed  in  the related work section.
  Afterwards,   pre-trained   language   models   on   large  corpus  like  BERT
  \cite{devlin2018bert}  and  RoBERTa \cite{liu2019roberta} has been proven more
  powerful  in  solving  many  NLP  tasks  including  short-text  classification
  \cite{madabushi2020cost}.      Especially      for      few-shot     scenarios
  \cite{yu2018diverse},     pre-trained     model    based    on    transformers
  \cite{vaswani2017attention}, shown in Fig. \ref{Fig.1.sub.1}, tends to do more
  help to reducing the negative effects brought by scarcity of training data.

  The  second  one,  \emph{similarity model} is also called sentence-pair model.
  Its  motivation  originates  from  the idea of Information Retrival (IR) based
  chatbot   \cite{jafarpour2010filter,   leuski2011npceditor}.   Having   a  Q-A
  (Question-Answer)  pairs dataset and user query Q, the IR based conversational
  system will look up in the Q-A dataset for the pair (Q', A') that best matches
  query  Q  through  semantic  analysis  and  returns  A'  as  the  answer  to Q
  \cite{mnasri2019recent}. In this way, sentence-pair model pre-trained on large
  corpus  of  semantic  similarity  identification task can be applied for being
  used  as  a  tool  to  identify  the class with highest semantic similarity to
  customer's    query.    A    common    model    structure    of    pre-trained
  \cite{devlin2018bert} sentence-pair model is based on multiple cross-attention
  mechanism  \cite{barkan2020scalable},  shown in Fig. \ref{Fig.1.sub.2}. Due to
  the   fact   that   many   experiment   results   have   shown   that  RoBERTa
  \cite{liu2019roberta} is an improved version of BERT, and has achieved amazing
  results  on  both  text  classification and sentence-pairs semantic similarity
  tasks, we choose RoBERTa as an important baseline approach in this paper.

@@

  The  above limitation motivates us to propose a joint system that is capble of
  taking  advantagges  from  both  classification model and similarity model. We
  calle  it  SFC,  short  for  similairty model fused with classification Model,
  shown  in  Fig.  \ref{fig:ver.1.3}.  To  better train SFC, we further bring in
  multi-task   learning   \cite{caruana1993multitask,collobert2008unified}.  Our
  basic  idea  cames  from  two  stages. In the first stage, we use an auxiliary
  model  to  select top-K most possible labels for a input. This model can be an
  elastic search  \cite{divya2013elasticsearch}  or  a  text classification model
  trained on current task-specific chatbot data. In the second stage, we build a
  classification  model  whose  main modules are similarity models. Then, this
  structure  derives two goals to train towards, the classification loss for our
  application,  and the similarity loss for the main modules. When the two stages
  are  independently  optimized,  we call \emph{two-stage SFC}. We further find,
  the  the  quality  of  output  from  the  first  stage  might  limit the final
  performance  of  system,  since  the  candidate  labels for each user query is
  always  fixed  in  the  whole  multi-task  training  process. This observation
  motivates  us  to  further  improve  the  two-stage SFC into a joint training,
  called  \emph{joint  SFC}. In this way, the sentence pairs associated with the
  top-K  classes  also  become dynamic, and the text classification model in the
  first  stage  will also be further optimized to provide better top-K result by
  the final training loss.

@@

  The  second  step  is  \emph{negative sentence pair sampling}: As for negative
  sentence  pairs,  we will also randomly sample $P$ sentences for each class in
  the negative candidate class label set $\mathcal{C'}\backslash \{y_{i}\}$. The
  class  labels  in  $\mathcal{C'}\backslash  \{y_{i}\}$  are  the  set  of most
  confusing  class  labels comparing to the ground truth ${y_{i}}$, so we assume
  that  the  sentence  pairs  with  negative  label  grouped by user input query
  $x_{i}$  and sentences with class labels in $\mathcal{C'}\backslash \{y_{i}\}$
  are  strong  adversarial  negative  samples for sentence-pair similarity model
  that  can  help  enhance  the training speed and performance. According to the
  same  method  as  positive  sentence  pair  sampling, we can obtain the set of
  sentence  pairs  with  negative label as $\mathcal{N'}_{i}=\{[(x_{i}, x'_{1}),
  0], ..., [(x_{i}, x'_{P\cdot K}), 0]\}$

@@

  We  already  know  that the size of ${\Phi}^S$ in task 1 is $M\times 2=(N\cdot
  K\cdot  P)\times  2$, where $N$ is the total number of data points in original
  training  data $\mathcal{D}$, $K$ is a hyperparameter that controls the number
  of   candidate   class  labels  provided  by  stage  1,  and  $P$  is  also  a
  hyperparameter that controls the number of sentences we should randomly sample
  from each candidate class labels. Now, for the average pooling layer, we first
  reshape  ${\Phi}^S$  into  the  size  of $N\times K\times P\times 2$, and then
  split  it  into  ${\Phi}^{K,pos}$  and ${\Phi}^{K,neg}$, and both of them will
  have  the  size  of  $N\times  K\times  P$.  In this way, ${\Phi}^{K,pos}$ can
  represent  the  level  of  similarity  for each sentence pair, and in the mean
  time,  ${\Phi}^{K,neg}$  can  represent  the  level  of dissimilarity for each
  sentence  pair. Then, we can do average pooling for each candidate class among
  the  top-K  classes  as  shown  in  Equ.  \ref{eq:average  pooling:1} and Equ.
  \ref{eq:average pooling:2}.

  \begin{align}
    {\xi}_{i,j}^{K,pos} = \sum_{p=1}^{P}{\varPhi}_{i,j,p}^{K,pos} \label{eq:average pooling:1}\\
    {\xi}_{i,j}^{K,neg} = \sum_{p=1}^{P}{\varPhi}_{i,j,p}^{K,neg}
    \label{eq:average pooling:2}
  \end{align}

@@

          RoBERTa-large & \multirow{2}*{$\textbf{0.8974i}$} &

@@

          \multirow{2}*{$\textbf{0.9099}$} & \multirow{2}*{$i\textbf{0.9227}$} & 

@@

      test  set to be fixed. The highest scores for each data setting are marked

@@

  \begin{enumerate}
    \item \emph{Pretrained model based classification system}, which consists of
    BERT,  RoBERTa,  and  ALBERT  \cite{lan2019albert}  based.  These pretrained
    models  are  practically  proven  to  achieve  outstanding  performances  in
    classification task and other NLP tasks.

    \item   \emph{Non-pretrained   model  based  classification  system},  which
    consists  of  a typical CNN based TextCNN \cite{kim2014convolutional}, and a
    label-embedding based LEAM \cite{wang2018joint}. Regarding TextCNN, the
    model is based on output from RoBERTa tokenization, and the kernels are set
    as 1, 2, 3, 4, 5. Regarding LEAM, a literal class label is required, which
    is available only in HWU64, CLINC150, BANKING77. Thus, for other two
    datasets, we have to use its class number instead.  Empircally,
    a  non-pretrained  model  based  system  are  not  performing  as  well as a
    pretrained  models  based,  if  both is applicable in some task. Our listing
    these  non-pretrained model based systems here is to provide a comprehensive
    performance comparison on our datasets.

    \item   \emph{Pre-trained   model   based   similarity   model}.  We  choose
    RoBERTa-large  as  our  fundermental  module as our SFC systems are based on
    RoBERTa  too, and RoBERTa is empircally more effective than other pretrained
    models  in  the  short  text  classification  task  in  our  experiments. In
    inference,  we  use  an  elastic search to find a set of potential candidate
    labels for the similarity model, to guarantee a reasonable running time.
  \end{enumerate}

>>>>

  The   first  one,  \emph{text  classification},  includes  classic  long  text
  classification  and  short  text  classification  in  this  scenario.  Besides
  traditional  machine  learning  models  like  SVM \cite{suykens1999least},  boosting  tree \cite{tu2005probabilistic},  many work
  \cite{wen2016network}  choose  neural  network  such  as  convolutional neural
  networks (CNNs) \cite{kim2014convolutional,zhang2015character,conneau2016very}
  and       long       short       term       memory       networks      (LSTMs)
  \cite{mousa2017contextual,liu2016recurrent}   to   accomplish   the   task  of
  extracting  semantic  feature from limited amount of words. Their common model
  structure  is  adding  a  softmax classifier as the final layer and mapping to
  labels.  Another  two  interesting lines of work, label-word joint models, and
  joint  NER  and  classification  are  discussed  in  the related work section.
  Afterwards,   pre-trained   language   models   on   large  corpus  like  BERT
  \cite{devlin2018bert}  and  RoBERTa \cite{liu2019roberta} has been proven more
  powerful  in  solving  many  NLP  tasks  including  short-text  classification
  \cite{madabushi2020cost}.      Especially      for      few-shot     scenarios
  \cite{yu2018diverse},     pre-trained     model    based    on    transformers
  \cite{vaswani2017attention} tends to do more
  help to reducing the negative effects brought by scarcity of training data.

  The  second  one,  \emph{similarity model} is also called sentence-pair model.
  Its  motivation  originates  from  the idea of Information Retrival (IR) based
  chatbot   \cite{jafarpour2010filter,   leuski2011npceditor}.   Having   a  Q-A
  (Question-Answer)  pairs dataset and user query Q, the IR based conversational
  system will look up in the Q-A dataset for the pair (Q', A') that best matches
  query  Q  through  semantic  analysis  and  returns  A'  as  the  answer  to Q
  \cite{mnasri2019recent}. In this way, sentence-pair model pre-trained on large
  corpus  of  semantic  similarity  identification task can be applied for being
  used  as  a  tool  to  identify  the class with highest semantic similarity to
  customer's    query.    A    common    model    structure    of    pre-trained sentence-pair model is based on multiple cross-attention
  mechanism  \cite{barkan2020scalable}. Due to
  the   fact   that   many   experiment   results   have   shown   that  RoBERTa
  \cite{liu2019roberta} is an improved version of BERT \cite{devlin2018bert}, and has achieved amazing
  results  on  both  text  classification and sentence-pairs semantic similarity
  tasks, we choose RoBERTa as an important baseline approach in this paper.

@@

  The  above limitation motivates us to propose a joint system that is capble of
  taking  advantagges  from  both  classification model and similarity model. We
  calle  it  SFC,  short  for  similairty model fused with classification Model,
  shown  in  Fig.  \ref{fig:framework}.  To  better train SFC, we further bring in
  multi-task   learning   \cite{caruana1993multitask,collobert2008unified, liu2019multi}.  Our
  basic  idea  cames  from  two  stages. In the first stage, we use an auxiliary
  model  to  select top-K most possible labels for a input. This model can be an
  elastic search  \cite{divya2013elasticsearch}  or  a  text classification model
  trained on current task-specific chatbot data. In the second stage, we build a
  classification  model  whose  main modules are similarity models. Then, this
  structure  derives two goals to train towards, the classification loss for our
  application,  and the similarity loss for the main modules. When the two stages
  are  independently  optimized,  we call \emph{two-stage SFC}. We further find,
  the  the  quality  of  output  from  the  first  stage  might  limit the final
  performance  of  system,  since  the  candidate  labels for each user query is
  always  fixed  in  the  whole  multi-task  training  process. This observation
  motivates  us  to  further  improve  the  two-stage SFC into a joint training,
  called  \emph{joint  SFC}. In this way, the sentence pairs associated with the
  top-K  classes  also  become dynamic, and the text classification model in the
  first  stage  will also be further optimized to provide better top-K result by
  the final training loss.

@@

  The  second  step  is  \emph{negative sentence pair sampling}: As for negative
  sentence  pairs,  we will also randomly sample $P$ sentences for each class in
  the negative candidate class label set $\mathcal{C'}\backslash \{y_{i}\}$. The
  class  labels  in  $\mathcal{C'}\backslash  \{y_{i}\}$  are  the  set  of most
  confusing  class  labels comparing to the ground truth ${y_{i}}$, so we assume
  that  the  sentence  pairs  with  negative  label  grouped by user input query
  $x_{i}$  and sentences with class labels in $\mathcal{C'}\backslash \{y_{i}\}$
  are  strong  adversarial  negative  samples for sentence-pair similarity model
  that  can  help  enhance  the training speed and performance. According to the
  same  method  as  positive  sentence  pair  sampling, we can obtain the set of
  sentence  pairs  with  negative label as $\mathcal{N'}_{i}=\{[(x_{i}, x'_{1}),
  0], ..., [(x_{i}, x'_{P\cdot (K-1)}), 0]\}$

@@

  We  already  know  that the size of ${\Phi}^S$ in task 1 is $M\times 2=(N\cdot
  K\cdot  P)\times  2$, where $N$ is the total number of data points in original
  training  data $\mathcal{D}$, $K$ is a hyperparameter that controls the number
  of   candidate   class  labels  provided  by  stage  1,  and  $P$  is  also  a
  hyperparameter that controls the number of sentences we should randomly sample
  from each candidate class labels. Now, for the average pooling layer, we first
  reshape  ${\Phi}^S$  into  the  size  of $N\times K\times P\times 2$, and then
  split  it  into  ${\Phi}^{K,pos}$  and ${\Phi}^{K,neg}$, and both of them will
  have  the  size  of  $N\times  K\times  P$.  In this way, ${\Phi}^{K,pos}$ can
  represent  the  level  of  similarity  for each sentence pair, and in the mean
  time,  ${\Phi}^{K,neg}$  can  represent  the  level  of dissimilarity for each
  sentence  pair. Then, we can do average pooling for each candidate class among
  the  top-K  classes  as  shown  in  Equ.  \ref{eq:average  pooling}.

  \begin{align}
    {\xi}_{i,j}^{K,pos} = \sum_{p=1}^{P}{\varPhi}_{i,j,p}^{K,pos} \ \ \ \ \ \ \ 
    {\xi}_{i,j}^{K,neg} = \sum_{p=1}^{P}{\varPhi}_{i,j,p}^{K,neg}
    \label{eq:average pooling}
  \end{align}

@@

          RoBERTa-large & \multirow{2}*{$\textbf{0.8974}$} &

@@

          \multirow{2}*{$\textbf{0.9099}$} & \multirow{2}*{$\textbf{0.9227}$} & 

@@

      test  set to be fixed. The highest scores among all the baseline models and SFC variants for each data setting are both marked

@@

  \begin{enumerate}
    \item \emph{Pretrained model based classification system}, which consists of
    BERT \cite{devlin2018bert},  RoBERTa \cite{liu2019roberta},  and  ALBERT  \cite{lan2019albert}  based.  These pretrained
    models  are  practically  proven  to  achieve  outstanding  performances  in
    classification task and other NLP tasks.

    \item   \emph{Non-pretrained   model  based  classification  system},  which
    consists  of  a typical CNN based TextCNN \cite{kim2014convolutional}, and a
    label-embedding based LEAM \cite{wang2018joint}. Regarding TextCNN, the
    model is based on output from RoBERTa tokenization, and the kernels are set
    as 1, 2, 3, 4, 5. Regarding LEAM, a literal class label is required, which
    is available only in HWU64, CLINC150, BANKING77. Thus, for other two
    datasets, we have to use its class number instead.  Empirically,
    a  non-pretrained  model  based  system  are  not  performing  as  well as a
    pretrained  models  based,  if  both is applicable in some task. Our listing
    these  non-pretrained model based systems here is to provide a comprehensive
    performance comparison on our datasets.

    \item   \emph{Pre-trained   model   based   similarity   model}.  We  choose
    RoBERTa-large  as  our  fundamental  module as our SFC systems are based on
    RoBERTa  too, and RoBERTa is empirically more effective than other pretrained
    models  in  the  short  text  classification  task  in  our  experiments. In
    inference,  we  use  an  elastic search to find a set of potential candidate
    labels for the similarity model, to guarantee a reasonable running time.
  \end{enumerate}
